---
title: Useful R Commands
---

```{r}
#| label: load-packages-data
#| include: false
 
library(coursekata)
dat <- read.csv("data/colleges.csv")
```

## Notebook 1: Basic R & Data Exploration

```{r}
#| label: setup
#| include: false

# Packages used in the examples below
library(coursekata)
library(dplyr)

# Load the dataset used in examples (update the path if needed)
dat <- read.csv("data/colleges.csv")
```

### Assignment Operator (`<-`)

The assignment operator `<-` is used to store a value in an object.

**Syntax**

``` r
object_name <- value
```

**Example**

```{r}
#| label: assignment-operator

# Store value 10 in x
x <- 10

# Print out value of x
x
```

### `head()`

The `head()` function displays the first 6 rows of a dataset.

**Syntax**

``` r
head(dataset_name)
```

**Example**

```{r}
#| label: head-example

head(dat)
```

### `dim()`

The `dim()` function displays the dimensions (rows × columns) of a dataset.

**Syntax**

``` r
dim(dataset_name)
```

**Example**

```{r}
#| label: dim-example

dim(dat)
```

The first number is the number of rows (horizontal), and the second is the number of columns (vertical).

### `select()`

The `select()` function is used to select only certain variables (columns) from a dataset.

**Syntax**

``` r
new_data <- select(dataset_name, col1, col2, ...)
```

**Example**

```{r}
#| label: select-example

example_dat <- select(dat, name, median_debt, ownership, admit_rate, hbcu)
```

### `subset()`

The `subset()` function filters a dataset to obtain certain observations (rows), based on conditions.

**Syntax**

``` r
subset(dataset_name, condition)
```

**Example**

```{r}
#| label: subset-example

subset(example_dat, hbcu == "Yes" & admit_rate < 40)
```

### Common Conditions

| Operator | Meaning                  | Example.         |
|----------|--------------------------|------------------|
| `==`     | Equals exactly           | `x == 10`        |
| `!=`     | Does not equal           | `x != 10`        |
| `<`      | Less than                | `x < 10`         |
| `>`      | Greater than             | `x > 10`         |
| `<=`     | Less than or equal to    | `x <= 10`        |
| `>=`     | Greater than or equal to | `x >= 10`        |
| `|`      | Logical OR               | `x < 5 | x > 10` |
| `&`      | Logical AND              | `x > 5 & x < 10` |

### `arrange()`

The `arrange()` function orders the rows in a dataset based on one or more variables.

**Syntax**

``` r
arrange(data_name, column_name)
```

**Example (ascending order)**

``` r
arrange(example_dat, admit_rate)
```

```{r}
#| label: arrange-ascending
#| echo: false
arrange(example_dat, admit_rate) |> head()
```

### `desc()`

The `desc()` function modifies `arrange()` to put data in descending order.

**Syntax**

``` r
arrange(dataset_name, desc(column_name))
```

**Example (descending order)**

``` r
arrange(example_dat, desc(admit_rate))
```

```{r}
#| label: arrange-descending
#| echo: false

arrange(example_dat, desc(admit_rate)) |> head()
```

### (`$`) operator

The `$` operator selects a single variable (column) from a dataset.

**Syntax**

``` r
dataset_name$column_name
```

### `table()`

The `table()` function displays frequency counts for the values of a categorical variable.

**Syntax**

``` r
table(dataset_name$column_name)
```

**Example**

```{r}
#| label: table-example

table(dat$highest_degree)
```

### `gf_histogram()`

The `gf_histogram()` function creates a histogram of a quantitative variable.

**Syntax**

``` r
gf_histogram(~variable_name, data = dataset_name)
```

**Example**

```{r}
#| label: gf-histogram-example

gf_histogram(~admit_rate, data = dat)
```

### `gf_bar()`

The `gf_bar()` function creates a bar plot of a categorical variable.

**Syntax**

``` r
gf_bar(~variable_name, data = dataset_name)
```

**Example**

```{r}
#| label: gf-bar-example

gf_bar(~highest_degree, data = dat)
```

### (`~`) Symbol

The `~` symbol is often used to separate the outcome variable ($y$) and predictor variable ($x$) in graphs and models:

``` r
outcome ~ predictor
```

### `gf_boxplot()`

The `gf_boxplot()` function creates boxplots.

**Syntax**

``` r
gf_boxplot(outcome ~ predictor, data = dataset_name)
```

**Example**

```{r}
#| label: gf-boxplot-example

gf_boxplot(admit_rate ~ highest_degree, data = dat)
```

-   `admit_rate ~ highest_degree` means that `admit_rate` is the outcome ($y$) variable and `highest_degree` is the predictor ($x$) variable.
-   The boxplot displays `admit_rate` on the $y-$axis and `highest_degree` on the $x-$axis.

## Notebook 2: Simple Linear Regression

### (`~`) Symbol

The `~` symbol is used to separate the outcome ($y$) variable and the predictor ($x$) variable in graphs and models.

**Syntax**

``` r
outcome ~ predictor
```

**Example**

```{r}
#| label: model-example

default_rate ~ pct_PELL
```

In this example, `default_rate` is the outcome ($y$) variable and `pct_PELL` is the predictor ($x$) variable.

### `gf_point()`

The `gf_point()` function creates a scatterplot.

**Syntax**

``` r
gf_point(outcome ~ predictor, data = dataset_name)
```

**Example**

```{r}
#| label: gf-point-example

gf_point(default_rate ~ pct_PELL, data = dat)
```

-   `default_rate ~ pct_PELL` specifies default rate on the $y$-axis and percent PELL on the $x$-axis
-   `data = dat` tells the function which dataset to use

### (`%>%`) Pipe Operator

The `%>%` operator pipes the result of one command into the next command. It is often used to layer models on top of graphs.

**Syntax**

``` r
command_1 %>% command_2
```

**Example**

```{r}
#| label: gf-point-lm-pipe

gf_point(default_rate ~ pct_PELL, data = dat) %>%
gf_lm(color = "orange")
```

-   The first command creates the scatterplot
-   The second command overlays a linear model on top of the plot

### `gf_lm()`

The `gf_lm()` function plots a linear model on an existing graph.

**Syntax**

``` r
gf_lm(color = "color_name")
```

**Example**

```{r}
gf_point(default_rate ~ pct_PELL, data = dat) %>% 
  gf_lm(color = "orange")
```

-   `color = "orange"` sets the color of the fitted line

### `lm()`

The `lm()` function fits a linear regression model.

**Syntax**

``` r
model_name <- lm(outcome ~ predictor, data = dataset_name)
```

**Example**

```{r}
#| label: lm-example

PELL_model <- lm(default_rate ~ pct_PELL, data = dat)
PELL_model
```

-   `default_rate ~ pct_PELL` specifies the outcome and predictor
-   `data = dat` indicates the dataset used
-   `PELL_model <-` stores the model in an object named `PELL_model`
-   Printing the model displays the estimated coefficients

### `summary()`

The `summary()` function displays detailed information about a fitted regression model.

**Syntax**

``` r
summary(model_name)
```

**Example**

```{r}
#| label: summary-example

summary(PELL_model)
```

-   The output includes coefficient estimates and model statistics
-   Look for **Multiple R-squared** to find the $R^2$ value

## Notebook 3: Multiple Regression

### `lm()`

The `lm()` function can be used to fit a multiple regression model, where one outcome ($y$) variable is predicted using two or more predictors ($x_1, x_2, x_3, \ldots$).

**Syntax**

``` r
model_name <- lm(y ~ x1 + x2 + x3 + ..., data = dataset_name)
```

**Example**

```{r}
#| label: multiple-regression-example

tuition_grad_model <- lm(default_rate ~ net_tuition + grad_rate, data = dat)
tuition_grad_model
```

-   `default_rate` is the outcome ($y$) variable
-   `net_tuition` and `grad_rate` are predictor variables
-   The `+` symbol adds predictors to the model
-   `data = dat` tells R which dataset to use
-   `tuition_grad_model <-` stores the model in an object named `tuition_grad_model`

Running `tuition_grad_model` prints the estimated coefficients of the regression model.

### `summary()`

The `summary()` function displays detailed information about a multiple regression model.

**Syntax**

``` r
summary(model_name)
```

**Example**

```{r}
#| label: multiple-regression-summary

summary(tuition_grad_model)
```

-   Look for **Multiple R-squared** to find the $R^2$ value
-   Look for **Adjusted R-squared** to see the $R^2$ adjusted for the number of predictors

## Notebook 4: Machine Learning

### `poly()`

The `poly()` function adds polynomial terms to a regression model.

**Syntax**

``` r
poly(variable_name, degree)
```

**Example**

```{r}
#| label: sample_dat-example
#| include: false
dat <- read.csv("data/four_year_colleges.csv")
set.seed(2)
sample_dat <- sample(dat, size = 20)
```

```{r}
#| label: polynomial-model-example
sat_model_2 <- lm(default_rate ~ poly(SAT_avg, 2), data = sample_dat)

sat_model_2
```

-   `default_rate` is the outcome ($y$) variable
-   `SAT_avg` is the predictor ($x$) variable
-   `poly(SAT_avg, 2)` adds two polynomial terms: $x$ and $x^2$
-   `data = sample_dat` specifies the dataset used
-   `sat_model_2 <-` stores the fitted model in an object named `sat_model_2`

Running `sat_model_2` prints the estimated coefficients of the polynomial regression model.

### `predict()`

The `predict()` function generates predictions from a previously fitted model using new data.

**Syntax**

``` r
predict(model_name, newdata = dataset_name)
```

**Example**

``` r
predict(sat_model_2, newdata = test)
```

-   `sat_model_2` is a previously fitted model
-   `newdata = test` specifies the dataset on which predictions are made
-   The output gives predicted values based on the model \`\`\`
